---
title: "Spatiotemporal cross-validation"
author: "Dmitry Pavlyuk"
date: "June 4, 2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

This markdown document reproduces the research "Spatiotemporal cross-validation of urban traffic forecasting models", submitted to the 23rd EURO Working Group on Transportation Meeting, EWGT 2020.


```{r child='sampling.Rmd'}
```


Load necessary libraries
----------------
```{r libs}
library(needs)
needs(knitr)
needs(tidyverse)
needs(reshape2)
needs(ggplot2)
needs(geosphere)
needs(igraph)
needs(Metrics)
needs(imputeTS)
needs(e1071)
needs(MTS)
needs(forecast)
needs(doParallel)
needs(glasso)
needs(scales)
needs(stringr)
needs(e1071 )
needs(randomForest )
needs(Boruta)
needs(matrixStats)
needs(fsMTS)
needs(plot.matrix)
needs(lubridate)

source(file.path("R","prepare_data_functions.R"))
source(file.path("R","cv_functions.R"))
source(file.path("R","cv_utils.R"))
source(file.path("R","models_util.R"))
source(file.path("R","models_baseline.R"))
source(file.path("R","models_spvar.R"))

# Define files for intermediate results
dir<-"EWGT2020"
models.estimated.file <- file.path(dir,"models.rds")
results.file <- file.path(dir,"results.rds")
complete.rds <- file.path("data","prepared","ALL5-original.rds")
sample.rds <- file.path(dir,"data.rds")
ta<-5
options(pillar.sigfig = 4)
```

Prepare the research sample (or load if already prepared)
```{r sampling}
if (!file.exists(sample.rds)){
  mysample <- readRDS(complete.rds)
  colnames(mysample$data)<-gsub('.volume','',colnames(mysample$data))
  series<-colnames(mysample$data)[-1]
  central.node<-"rnd_90797"
  center<-mysample$config.nodes%>%filter(node_name==central.node)%>%select(node_lon,node_lat)%>%as.list
  nodes<-mysample$config.nodes%>%filter(node_name %in% series)%>%rowwise()%>%
    filter(distHaversine(c(node_lon,node_lat),c(center$node_lon,center$node_lat))<5000)
  write.csv(nodes,file.path(dir,"nodes-all.csv"), row.names = F)
  series <- nodes%>%select(node_name)%>%pull
  sampl<-list(data = mysample$data[,c("datetime",series)], shortest.distances=mysample$shortest.distances[series,series])
  saveRDS(sampl, file=sample.rds)
}else{
  sampl <- readRDS(sample.rds)
}
series<-colnames(sampl$data)[-1]
```

Define folders for feature sets
```{r define_folders}
maxlag<-3
distance.global <- file.path(dir,"fs",length(series),"distance",20160,maxlag,"2017-10-08_000000.rds")
CCF.global <- file.path(dir,"fs",length(series),"CCF",20160,maxlag,"2017-10-08_000000.rds")
ownlags.global <- file.path(dir,"fs",length(series),"ownlags",20160,maxlag,"2017-10-08_000000.rds")
```

Prepare and store feature sets
```{r fs}
estimateFeatures(fs.function=fsMTS::fsMTS, 
                 data = sampl$data, 
                 seriesNames = series,
                 trainingWindowSize = nrow(sampl$data),
                 forecastEvery = 1,
                 max.lag = 3,
                 fs.folder = file.path(dir,"fs",length(series),"CCF"),
                 req.packages = c("fsMTS"),
                 method="CCF",
                 shortest = sampl$shortest.distances,
                 step = ta,
                 clusterNumber=1
                 )
estimateFeatures(fs.function=fsMTS::fsMTS, 
                 data = sampl$data, 
                 seriesNames = series,
                 trainingWindowSize = nrow(sampl$data),
                 forecastEvery = 1,
                 max.lag = 3,
                 fs.folder = file.path(dir,"fs",length(series),"distance"),
                 req.packages = c("fsMTS"),
                 method="distance",
                 shortest = sampl$shortest.distances,
                 step = ta,
                 clusterNumber=1
                 )
estimateFeatures(fs.function=fsMTS::fsMTS, 
                 data = sampl$data, 
                 seriesNames = series,
                 trainingWindowSize = nrow(sampl$data),
                 forecastEvery = 1,
                 max.lag = 3,
                 fs.folder = file.path(dir,"fs",length(series),"ownlags"),
                 req.packages = c("fsMTS"),
                 method="ownlags",
                 step = ta,
                 clusterNumber=1
                 )

```


Estimate similarity of feature sets
```{r similarity_matrix}
mlist<-list(
  CCF=readRDS(CCF.global),
  Distance=readRDS(distance.global),
  ownlags=readRDS(ownlags.global)
)
msimilarity<-fsSimilarityMatrix(mlist, threshold = 0.1, method="Kuncheva")

needs(heatmaply)
plot(msimilarity, digits=2, col=rev(Spectral(21)),breaks=seq(-1,1,by=0.1), key=NULL, main="Sample-wise", cex.axis=0.7,ylab="",xlab="")
```

Forecasting
```{r models}
RunModels <- function(data, series, ta, forecastingSteps, forecastEvery,
                      validationSize, validationEnd, 
                      lagMatrix,shortestA, cvgrid, results.folder){
  dir.create(results.folder, recursive = T)
  start.time <- Sys.time()
  results<-tibble()
  models.estimated <- c()
  if (file.exists(models.estimated.file)) models.estimated <- readRDS(models.estimated.file)
  else saveRDS(models.estimated, models.estimated.file)
  if (file.exists(results.file)) results <- readRDS(results.file)
  else saveRDS(results, results.file)
  for (r in 1:nrow(cvgrid)){
    cv <- cvgrid[r,]
    print(paste("Tuned parameter's set",r,"of",nrow(cvgrid)))
    print(cv)
    validationStart <- validationEnd - validationSize
    trainingSize <- cv$trainingMinutes/ta
    dat.restricted<-data[(validationStart-trainingSize+1):validationEnd,]
    base_params<-list(data=dat.restricted, seriesNames=series,
                      forecastingSteps=forecastingSteps, forecastEvery=forecastEvery)
    params<-c(base_params,list(trainingWindowSize=trainingSize))
    
    if (length(forecastingSteps)>1) params$forecastingSteps=max(forecastingSteps)
    estimate.SpVAR(params,cv,results.folder)
    estimate.KNN(params,cv,results.folder)
    print(Sys.time() - start.time)
  }
}
```

FS -experiments
```{r limited_fs}
fs<-readRDS(CCF.global)
fs<-fsMTS::cutoff(fs,0.1)
dim(fs)
cs<-colSums(fs)
cs
for (i in 1:length(cs)){
  if (cs[i]==0){
    node<-names(cs)[i]
    dep<-paste0(node,c(".l1",".l2",".l3"))
    fs[dep,node]<-1
  }
}
colSums(fs)
sum(colSums(fs)==0)

rownames(fs)
res<-list()
exclude<-c()
level<-1
while(length(exclude)<ncol(fs)){
  level.exclude<-exclude
  elems<-colnames(fs)
  elems<- elems[!elems %in% exclude]
  eff<-list()
  for(node in elems){
    dep<-paste0(node,c(".l1",".l2",".l3"))
    depnum<-sum(fs[dep,!(colnames(fs) %in% level.exclude)])
    if (depnum==0){
      exclude<-append(exclude,node)
      res[[length(res)+1]]<-list(node=node, level=level)
    } else{
      eff[[length(eff)+1]]<-list(node=node, eff=depnum)
    }
  }
  if (length(level.exclude)==length(exclude)){#Nothing to remove
    df.eff<-as_tibble(bind_rows(eff))
    excl<-df.eff%>%arrange(eff)%>%first%>%select(node)%>%pull
    print(excl)
    exclude<-append(exclude,excl)
    res[[length(res)+1]]<-list(node=excl, level=level)
  }
  level<-level+1
}
node.levels<-bind_rows(res)
fs.dir<-file.path(dir,"fs","CCF")
dir.create(fs.dir)

node.levels<-node.levels%>%arrange(level)

wfs<-fs
excluded <-0
for (i in 1:nrow(node.levels)){
  fs2<-wfs
  node <- node.levels[i,]$node
  dep<-paste0(node,c(".l1",".l2",".l3"))
  fs2[dep,]<-0
  if (sum(colSums(fs2)==0)==0){
    wfs<-fs2
    excluded<-excluded+1
    if (excluded %% 10 ==0){
      print(sum(wfs))
      saveRDS(wfs,file.path(fs.dir,paste0("minus",excluded,".rds")))
    }
  }
}
colSums(fs)
colSums(wfs)



```

Forecasting - experiments
```{r experiments}

distance.minus10<-file.path(dir,"fs","distance","minus10.rds")
distance.minus30<-file.path(dir,"fs","distance","minus30.rds")
distance.minus60<-file.path(dir,"fs","distance","minus60.rds")
distance.minus80<-file.path(dir,"fs","distance","minus80.rds")
CCF.minus10<-file.path(dir,"fs","CCF","minus10.rds")
CCF.minus30<-file.path(dir,"fs","CCF","minus30.rds")
CCF.minus60<-file.path(dir,"fs","CCF","minus60.rds")
CCF.minus80<-file.path(dir,"fs","CCF","minus80.rds")

dat<-as.data.frame(sampl$data)
rownames(dat)<-format(dat$datetime,"%Y-%m-%d %H:%M:%S")

week<-24*60*7
data<-sampl$data
validationSize <- nrow(data)-week/ta
validationEnd <- nrow(data)
forecastEvery<-12
forecastingSteps<-12

# Define a grid for tuning parameters
cvgrid <- expand.grid(trainingMinutes = c(week),
                      include.mean = c(F),
                      arLags = c(maxlag),
                      stationary =  c(T),
                      allowdrift =  c(F),
                      fs.folder=c(CCF.minus10, distance.minus10,
                                  CCF.minus30, distance.minus30,
                                  CCF.minus60, distance.minus60),
                      threshold=c(1),
                      ownlags=c(F),
                      kneighbours=c(36))
source(file.path("R","models_spvar.R"))
source(file.path("R","cv_functions.R"))
# Run all models
RunModels(data, series, ta, forecastingSteps, forecastEvery,
                      validationSize, validationEnd, 
                      lagMatrix, shortestA, cvgrid,
                      results.folder=file.path(dir,"models"))
 
```


```{r analyse_results}
#params <- c('trainingMinutes','allowdrift','stationary','arLags','include.mean','fs.folder','threshold','ownlags')
# models.estimated <- readRDS(file.path(dir,"baseline.models.rds"))
# results <- readRDS(results.file)
# models.estimated <- readRDS(models.estimated.file)
# results<-results%>%mutate(fs.folder=gsub(file.path(dir,"fs",64),"",fs.folder))
# summary(results$actual)
# #results<-results%>%filter(abs(actual)>100)
# #results%>%select(datetime,detector,actual,forecasted)
# actu<-mysample$data.orig%>%gather(key="detector", value="actual.orig", -one_of("datetime"))%>%
#   mutate(detector=gsub(".volume","",detector))
# r<-results%>%left_join(actu, by=c("datetime"="datetime","detector"="detector"))
# r1<-r%>%mutate(actual=actual+actual.orig,forecasted=forecasted+actual.orig)

reslist<-list()
results<-NULL
fold<-file.path(dir,"models")
for (f in list.files(fold)){
  print(f)
  reslist[[length(reslist)+1]]<-readRDS(file.path(fold, f))
}
results<-bind_rows(reslist)
params <- c('trainingMinutes','arLags','fs.folder','threshold','ownlags')
s<-cvSummary(results, rmse, cumulative=F, params=params)%>%
    mutate(modelid=paste(model,trainingMinutes, arLags,fs.folder,threshold,ownlags))%>%
    mutate(modelid=gsub(" NA","",modelid))
s%>%ungroup%>%mutate(key=paste(model,trainingMinutes,fs.folder))%>%select(key,threshold,forecast_horizon,RMSE)%>%spread(key=forecast_horizon, value=RMSE)%>%print(n=34)

```