---
title: "Ensemble feature selection"
author: "Dmitry Pavlyuk"
date: "Nov 09, 2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

This markdown document reproduces the research "Robust Learning of Spatiotemporal Urban Traffic Flow Relationships" (to be submitted)


```{r child='sampling.Rmd'}
```


Load necessary libraries
----------------
```{r libs}
library(needs)
needs(knitr)
needs(tidyverse)
needs(reshape2)
needs(ggplot2)
needs(geosphere)
needs(igraph)
needs(Metrics)
needs(imputeTS)
needs(e1071)
needs(MTS)
needs(forecast)
needs(doParallel)
needs(glasso)
needs(scales)
needs(stringr)
needs(e1071 )
needs(randomForest )
needs(Boruta)
needs(matrixStats)
needs(fsMTS)
needs(plot.matrix)
needs(lubridate)

source(file.path("R","prepare_data_functions.R"))
source(file.path("R","cv_functions.R"))
source(file.path("R","cv_utils.R"))
source(file.path("R","models_util.R"))
source(file.path("R","models_baseline.R"))
source(file.path("R","models_spvar.R"))

# Define files for intermediate results
dir<-"RobustFS2020"
models.estimated.file <- file.path(dir,"models.rds")
results.file <- file.path(dir,"results.rds")
complete.rds <- file.path("data","prepared","ALL5-original.rds")
sample.rds <- file.path(dir,"data64.rds")
ta<-5
options(pillar.sigfig = 4)
```

Prepare the research sample (or load if already prepared)
```{r sampling}
if (!file.exists(sample.rds)){
  mysample <- readRDS(complete.rds)
  colnames(mysample$data)<-gsub('.volume','',colnames(mysample$data))
  series<-colnames(mysample$data)[-1]
  central.node<-"rnd_90797"
  center<-mysample$config.nodes%>%filter(node_name==central.node)%>%select(node_lon,node_lat)%>%as.list
  nodes<-mysample$config.nodes%>%filter(node_name %in% series)%>%rowwise()%>%
    filter(distHaversine(c(node_lon,node_lat),c(center$node_lon,center$node_lat))<25000)
  write.csv(nodes,file.path(dir,"nodes-all.csv"), row.names = F)
  series <- nodes%>%select(node_name)%>%pull
  morning.rush<-mysample$data%>%filter(hour(datetime)>=7 & hour(datetime)<10 & !(weekdays(datetime)%in%c("Sunday", "Saturday")))
  evening.rush<-mysample$data%>%filter(hour(datetime)>=17 & hour(datetime)<20 & !(weekdays(datetime)%in%c("Sunday", "Saturday")))
  selectMostImportant <- function(data, series, rho, topn, shortest, distance.lim=4){
    m.cor<-stats::cor(data[,series])
    m.cor[is.na(m.cor)]<-0
    gl.m <- glasso::glasso(m.cor,rho=rho)
    rs.m<-as_tibble(rowSums(abs(gl.m$wi)>0))%>%mutate(node=series)
    hist(rs.m%>%select(value)%>%pull)
    s.m<-rs.m%>%top_n(topn, wt=value)%>%arrange(desc(value))%>%select(node)%>%pull
    s.m.filtered <- c()
    for (s in s.m){
      ser<-c(s,s.m.filtered)
      m<-shortest[ser,ser]
      if (min(m+(distance.lim+0.01)*diag(ncol(m)))>=distance.lim){
        s.m.filtered<-append(s.m.filtered,s)
      }
    }
    return(s.m.filtered)
  }
  morning.nodes <- selectMostImportant(morning.rush, series = c(series), 
                    rho=0.3, topn = 50, shortest=mysample$shortest.distances,distance.lim=1)
  length(morning.nodes)
  write.csv(mysample$config.nodes%>%filter(node_name %in% morning.nodes),file.path(dir,"nodes-MR50.csv"),  row.names = F)
  evening.nodes <- selectMostImportant(evening.rush, series = c(series), rho=0.3, topn = 50, shortest=mysample$shortest.distances,distance.lim=1)
  length(evening.nodes)
  write.csv(mysample$config.nodes%>%filter(node_name %in% evening.nodes),file.path(dir,"nodes-ER50.csv"),  row.names = F)
  write.csv(mysample$config.nodes%>%
              filter(node_name %in% gsub('.volume','',intersect(morning.nodes,evening.nodes))),file.path(dir,"nodes-ER+MR.csv"),  row.names = F)
  intersect(morning.nodes,evening.nodes)
  length(union(morning.nodes,evening.nodes))
  series <- union(morning.nodes,evening.nodes)
  dat<-mysample$data.orig%>%gather(key = "node", value="value", -datetime)%>%
    filter(node %in% paste0(series,".volume"))%>%mutate(node=gsub(".volume","",node), dt=as.POSIXct(format(datetime, "%H:%M"),format="%H:%M",tz="GMT"))%>%
    group_by(node,dt)%>%summarize(value=median(value))
  date.labels = seq(from = min(dat$dt),to = max(dat$dt), by = "hour")
  dat%>%filter(node %in% morning.nodes)%>%ggplot(aes(x = dt, y = value, col=node, group=node)) + geom_line()+ 
    theme(legend.position = "none")+scale_x_datetime(labels = date_format("%H:%M"), date_breaks = "4 hours")+ xlab("TIme of the day") + ylab("Traffic volume")
  dat%>%filter(node %in% evening.nodes)%>%ggplot(aes(x = dt, y = value, col=node, group=node))  + geom_line()+ 
    theme(legend.position = "none")+scale_x_datetime(labels = date_format("%H:%M"), date_breaks = "4 hours")+ xlab("TIme of the day") + ylab("Traffic volume")
  sampl<-list(data = mysample$data[,c("datetime",series)], shortest.distances=mysample$shortest.distances[series,series])
  saveRDS(sampl, file=sample.rds)
}else{
  sampl <- readRDS(sample.rds)
}
series<-colnames(sampl$data)[-1]
```

Define folders for feature sets
```{r define_folders}
maxlag<-3
CCF.folder <- file.path(dir,"fs",length(series),"CCF",36,maxlag)
MI.folder <- file.path(dir,"fs",length(series),"MI",36,maxlag)
GLASSO.folder <- file.path(dir,"fs",length(series),"GLASSOloc",36,maxlag,0.3)
LARS.folder <- file.path(dir,"fs",length(series),"LARS",36,maxlag)
RF.folder <- file.path(dir,"fs",length(series),"RF",36,maxlag)
EnsVoting.folder <- file.path(dir,"fs",length(series),"EnsVoting",36,maxlag,0.1)
EnsRanking.folder <- file.path(dir,"fs",length(series),"EnsRanking",36,maxlag,0.1)

CCFg.folder <- file.path(dir,"fs",length(series),"CCFg",36,maxlag)
MIg.folder <- file.path(dir,"fs",length(series),"MIg",36,maxlag)
GLASSOg.folder <- file.path(dir,"fs",length(series),"GLASSOlocg",36,maxlag,0.3)
LARSg.folder <- file.path(dir,"fs",length(series),"LARSg",36,maxlag)
RFg.folder <- file.path(dir,"fs",length(series),"RFg",36,maxlag)
EnsVotingg.folder <- file.path(dir,"fs",length(series),"EnsVotingg",36,maxlag,0.1)
EnsRankingg.folder <- file.path(dir,"fs",length(series),"EnsRankingg",36,maxlag,0.1)

distance.folder <- file.path(dir,"fs",length(series),"distance",20160,maxlag,"2017-10-08_000000.rds")
CCF.global <- file.path(dir,"fs",length(series),"CCF",20160,maxlag,"2017-10-08_000000.rds")
GLASSO.global <- file.path(dir,"fs",length(series),"GLASSOloc",20160,maxlag,0.3,"2017-10-08_000000.rds")
MI.global <- file.path(dir,"fs",length(series),"MI",4032,maxlag,"2017-08-13_000000.rds")
LARS.global <- file.path(dir,"fs",length(series),"LARS",20160,maxlag,"2017-10-08_000000.rds")
RF.global <- file.path(dir,"fs",length(series),"RF",20160,maxlag,"2017-10-08_000000.rds")
EnsVoting.global <- file.path(dir,"fs",length(series),"EnsVoting",20160,maxlag,0.1,"2017-10-08_000000.rds")
EnsRanking.global <- file.path(dir,"fs",length(series),"EnsRanking",20160,maxlag,0.1,"2017-10-08_000000.rds")


CCF.folder.drift <- file.path(dir,"fs",length(series),"CCFdrift",36,maxlag)
GLASSO.folder.drift <- file.path(dir,"fs",length(series),"GLASSOlocdrift",36,maxlag)
MI.folder.drift <- file.path(dir,"fs",length(series),"MIdrift",36,maxlag)
LARS.folder.drift <- file.path(dir,"fs",length(series),"LARSdrift",36,maxlag)
RF.folder.drift <- file.path(dir,"fs",length(series),"RFdrift",36,maxlag)
EnsVoting.folder.drift <- file.path(dir,"fs",length(series),"EnsVotingdrift",36,maxlag,0.1)
EnsRanking.folder.drift <- file.path(dir,"fs",length(series),"EnsRankingdrift",36,maxlag,0.1)
```

Prepare and store feature sets
```{r fs}
estimateFeatures(fs.function=fsMTS::fsMTS, 
                 data = sampl$data, 
                 seriesNames = series,
                 trainingWindowSize = nrow(sampl$data),#12*24*0.5,
                 forecastEvery = 1,
                 max.lag = 3,
                 fs.folder = file.path(dir,"fs",length(series),"GLASSOloc"),
                 req.packages = c("fsMTS"),
                 method="GLASSO",
                 rho=0.5,
                 shortest = shortestA,
                 step = ta,
                 localized = T,
                 show.progress=T,
                 clusterNumber=1
                 )

estimateFeaturesGlobal(fs.function=fsMTS::fsMTS, 
                 data = sampl$data, 
                 seriesNames = series,
                 trainingWindowSize = 36,
                 forecastEvery = 1,
                 max.lag = 3,
                 ta=ta,
                 fs.folder = file.path(dir,"fs",length(series),"MIg"),
                 req.packages = c("fsMTS","lubridate","tibble","tidyverse"),
                 method="MI",
                 rho=0.3,
                 shortest = shortestA,
                 req.functions=c("daySec"),
                 step = ta,
                 localized = F,
                 clusterNumber = 1
                 )
```

Prepare and store feature ensembles
```{r create_ensembles}
mlist <- list(Distance = fs.distance.global,
              CCF = fs.CCF.global,
              GLASSO = fs.GLASSO.global,
              MI = fs.MI.global,
              LARS = fs.LARS.global,
              RF = fs.RF.global)

th<-0.1
plot(fs.CCF.global, col=rev(heat.colors(10)), key=NULL, main="FS")
plot(cutoff(fs.CCF.global,th), col=rev(heat.colors(10)), key=NULL, main="FS")
plot(cutoff(fs.LARS.global,th), col=rev(heat.colors(10)), key=NULL, main="FS")
mlist[["EnsembleRank"]] <- fsMTS::fsEnsemble(mlist, threshold = th, method="ranking")
mlist[["EnsembleMajV"]] <- fsMTS::fsEnsemble(mlist, threshold = th, method="majority")
(msimilarity <- fsSimilarityMatrix(mlist,threshold = th, method="Hamming"))
colnames(mlist[["EnsembleRank"]])<-colnames(mlist[["EnsembleMajV"]])
rownames(mlist[["EnsembleRank"]])<-rownames(mlist[["EnsembleMajV"]])
  
plot(mlist[["EnsembleMajV"]], col=rev(heat.colors(10)), key=NULL, main="FS")
val<-fsSimilarity(mlist[["EnsembleRank"]],mlist[["EnsembleMajV"]], method="Kuncheva")
fsMTS::fsSparsity(mlist[["EnsembleRank"]])
fsMTS::fsSparsity(mlist[["EnsembleMajV"]])
dir.create(file.path(dir,"fs",length(series),"EnsVoting",20160,maxlag,th), recursive = T)
dir.create(file.path(dir,"fs",length(series),"EnsRanking",20160,maxlag,th), recursive = T)
saveRDS(mlist[["EnsembleMajV"]], file.path(dir,"fs",length(series),"EnsVoting",20160,maxlag,th,"2017-10-08_000000.rds"))
saveRDS(mlist[["EnsembleRank"]], file.path(dir,"fs",length(series),"EnsRanking",20160,maxlag,th,"2017-10-08_000000.rds"))

EnsVotingg.folder <- file.path(dir,"fs",length(series),"EnsVotingg",36,maxlag,th)
dir.create(EnsVotingg.folder, recursive = T)
EnsRankingg.folder <- file.path(dir,"fs",length(series),"EnsRankingg",36,maxlag,th)
dir.create(EnsRankingg.folder, recursive = T)

for (f in list.files(CCFg.folder)){
  fs.CCF.daytime <- readRDS(file.path(CCFg.folder,f))
  fs.GLASSO.daytime <- readRDS(file.path(GLASSOg.folder,f))
  fs.MI.daytime <- readRDS(file.path(MIg.folder,f))
  fs.LARS.daytime <- readRDS(file.path(LARSg.folder,f))
  fs.RF.daytime <- readRDS(file.path(RFg.folder,f))
  mlist <- list(Distance = fs.distance.global,
              CCF = fs.CCF.daytime,
              GLASSO = fs.GLASSO.daytime,
              MI = fs.MI.daytime,
              LARS = fs.LARS.daytime,
              RF = fs.RF.daytime)
  EnsembleRank <- fsMTS::fsEnsemble(mlist, threshold = th, method="ranking")
  EnsembleMajV <- fsMTS::fsEnsemble(mlist, threshold = th, method="majority")
  colnames(EnsembleRank)<-colnames(EnsembleMajV)
  rownames(EnsembleRank)<-rownames(EnsembleMajV)
  saveRDS(EnsembleMajV,file.path(EnsVotingg.folder,f))
  saveRDS(EnsembleRank,file.path(EnsRankingg.folder,f))
  print(paste(f,"Sparsity",fsMTS::fsSparsity(EnsembleMajV)))
}

EnsVoting.folder <- file.path(dir,"fs",length(series),"EnsVoting",36,maxlag,th)
dir.create(EnsVoting.folder, recursive = T)
EnsRanking.folder <- file.path(dir,"fs",length(series),"EnsRanking",36,maxlag,th)
dir.create(EnsRanking.folder, recursive = T)

for (f in list.files(CCF.folder)){
  fs.CCF.dynamic <- readRDS(file.path(CCF.folder,f))
  fs.GLASSO.dynamic <- readRDS(file.path(GLASSO.folder,f))
  fs.MI.dynamic <- readRDS(file.path(MI.folder,f))
  fs.LARS.dynamic <- readRDS(file.path(LARS.folder,f))
  fs.RF.dynamic <- readRDS(file.path(RF.folder,f))
  mlist <- list(Distance = fs.distance.global,
              CCF = fs.CCF.dynamic,
              GLASSO = fs.GLASSO.dynamic,
              MI = fs.MI.dynamic,
              LARS = fs.LARS.dynamic,
              RF = fs.RF.dynamic)
  EnsembleRank <- fsMTS::fsEnsemble(mlist, threshold = th, method="ranking")
  EnsembleMajV <- fsMTS::fsEnsemble(mlist, threshold = th, method="majority")
  colnames(EnsembleRank)<-colnames(EnsembleMajV)
  rownames(EnsembleRank)<-rownames(EnsembleMajV)
  saveRDS(EnsembleMajV,file.path(EnsVoting.folder,f))
  saveRDS(EnsembleRank,file.path(EnsRanking.folder,f))
  print(paste(f,"Sparsity",fsMTS::fsSparsity(EnsembleMajV)))
}
```

Prepare feature drifts
```{r create_drift}
EnsVoting.folder.drift <- file.path(dir,"fs",length(series),"EnsVotingdrift",36,maxlag,0.1)
EnsRanking.folder.drift <- file.path(dir,"fs",length(series),"EnsRankingdrift",36,maxlag,0.1)
SMOOTH_SIZE <- 6
prev <- list()
fold<-EnsRanking.folder
fold.drift <- EnsRanking.folder.drift
if (!file.exists(fold.drift)) dir.create(fold.drift, recursive = T)
p<-NULL
pOrig<-NULL
for (f in list.files(fold)){
  fs <- readRDS(file.path(fold,f))
  prev[[length(prev)+1]]<-fs
  if (length(prev)>SMOOTH_SIZE) prev[[1]] <- NULL
  if  (length(prev)==1){
    fs.drift <- fs
  }else{
    fs.drift <-apply(simplify2array(prev), 1:2, mean)
  }
  if (length(prev)>1) print(paste(f," ",fsSimilarity(p,fs.drift,cutoff=T, threshold = 0.1)," ",fsSimilarity(pOrig,fs,cutoff=T, threshold = 0.1)))
  p<-fs.drift
  pOrig<-fs
  saveRDS(fs.drift,file.path(fold.drift,f))
}

```

Estimate stability of feature sets
```{r one_step_similarity}
oneStepSimilarity <- function(folder, date.from,date.to, threshold, method){
  dates <- seq(from=date.from, to=date.to, by="5 min")
  res <- list()
  fs.prev<-NULL
  fmt <- "%Y-%m-%d_%H%M%S"
  for(row in 1:length(dates)){
    name<-format(dates[row],fmt)
    #filename <- paste0(name,".rds")
    filename <- file.path(paste0(wday(dates[row]),"_",daySec(dates[row]),".rds"))
    if (file.exists(file.path(folder, filename))){
      fs <- readRDS(file.path(folder, filename))
      if (!is.null(fs.prev)){
        res[[length(res)+1]] <- list(datetime=dates[row], 
                         similarity=fsMTS::fsSimilarity(fs.prev, fs,cutoff=T, 
                                                        threshold = threshold, method=method),
                         threshold = threshold,
                         method=method,
                         key=folder)
      }
      fs.prev <- fs
    }else{
      fs.prev <- NULL
    }
  }
  return(bind_rows(res))
}

res<-list()
for (f in list.files(file.path(dir,"onestep"))){
  res[[length(res)+1]]<-readRDS(file.path(dir,"onestep",f))
}
all<-bind_rows(res)
saveRDS(res, file.path(dir,"onestep","all.rds"))

res<-all%>%mutate(wd=wday(datetime))%>%filter(wd==3)%>%mutate(dow_time=format(datetime,"%H:%M"))

res%>%filter(method=="Kuncheva",fs=="RF", threshold>0)%>%mutate(key=paste(fs, strategy))%>%group_by(strategy,dow_time)%>%summarise(mean.similarity=mean(similarity),dt=first(datetime))%>%ggplot(aes(x = dt, y=mean.similarity, group=strategy, col=strategy))+geom_line(size=1)+scale_x_datetime(labels = date_format("%H:%M", tz="Europe/Helsinki"),date_breaks = "4 hours",expand = c(0,0))+ xlab("Time of the day") + ylab("One-step similarity")+ guides(col=guide_legend(title="Feature selection strategy"))+ theme(legend.position="bottom")+scale_color_brewer(palette="Dark2")

res%>%filter(method=="Kuncheva",strategy=="dynamic", threshold>0)%>%mutate(key=paste(fs, strategy))%>%group_by(fs,dow_time)%>%summarise(mean.similarity=mean(similarity),dt=first(datetime))%>%ggplot(aes(x = dt, y=mean.similarity, group=fs, col=fs))+geom_line(size=1)+scale_x_datetime(labels = date_format("%H:%M", tz="Europe/Helsinki"),date_breaks = "4 hours",expand = c(0,0))+ xlab("Time of the day") + ylab("One-step similarity")+ guides(col=guide_legend(title="Feature selection algorithm"))+ theme(legend.position="bottom")+scale_color_brewer(palette="Dark2")

res%>%group_by(fs,strategy)%>%summarise(mean.onestep=mean(similarity))%>%spread(strategy,"mean.onestep")
```

```{r daily_stability}
loadDay<-function(datetime, folder){
  fmt <- "%Y-%m-%d_%H%M%S"
  dates <- seq(from=datetime, to=datetime+(24*60-1)*60, by="5 min")
  mlist<-list()
  for(row in 1:length(dates)){
    name<-format(dates[row],fmt)
    #filename <- paste0(name,".rds")
    filename <- file.path(paste0(wday(dates[row]),"_",daySec(dates[row]),".rds"))
    if (file.exists(file.path(folder, filename))){
      fs <- readRDS(file.path(folder, filename))
      mlist[[name]]<-fs
    }
  }  
  return(mlist)
}
needs(lubridate)

fmt <- "%Y-%m-%d_%H%M%S"
date.from <- as.POSIXct("2017-07-31_000000",format=fmt)
date.to <- as.POSIXct("2017-08-06_000000",format=fmt)
days <- seq(from=date.from, to=date.to, by="1 days")
cl <- makeCluster(5, outfile="outfile.txt")
registerDoParallel(cl)
foreach (d = 1:length(days),.export=c("loadDay"),
                  .packages=c("fsMTS","dplyr","lubridate")) %dopar% {
  res<-list()
  day<-days[d]
  for (fold in c(CCFg.folder,MIg.folder,GLASSOg.folder,LARSg.folder,RFg.folder,EnsVotingg.folder,EnsRankingg.folder)){
    print(paste(day,fold))
    ml<-loadDay(day,fold)
    msim<-fsSimilarityMatrix(ml,threshold=0.1,method="Kuncheva")
    print(paste("Stability", mean(msim)))
    res[[length(res)+1]]<-list(folder=fold,date=day,stability=mean(msim))
  }
  result<-bind_rows(res)
  fmt <- "%Y-%m-%d_%H%M%S"
  filename<-paste0(format(day,fmt),".rds")
  print(filename)
  saveRDS(result, file.path("stability","daytime", filename))
}
res<-list()
for (f in list.files(file.path("stability","daytime"))){
  res[[length(res)+1]]<-readRDS(file.path("stability","daytime",f))
}
all<-bind_rows(res)
all%>%group_by(folder)%>%summarise(mean.stability = mean(stability))
plot(mlist[["EnsRanking"]], col=rev(heat.colors(10)), key=NULL, main="EnsRanking",ylab="",xlab="", axis.col = NULL, axis.row = NULL,border="lightgrey")
```



Estimate similarity of feature sets
```{r similarity_matrix}
mlist<-list(
  CCF=readRDS(CCF.global),
  GLASSO=readRDS(GLASSO.global),
  MI=readRDS(MI.global),
  LARS=readRDS(LARS.global),
  RF=readRDS(RF.global),
  Distance=readRDS(distance.folder),
  EnsVoting=readRDS(EnsVoting.global),
  EnsRanking=readRDS(EnsRanking.global)
)
plot(mlist[["EnsRanking"]], col=rev(heat.colors(10)), key=NULL, main="EnsRanking",ylab="",xlab="", axis.col = NULL, axis.row = NULL,border="lightgrey")
plot(mlist[["EnsVoting"]], col=rev(heat.colors(10)), key=NULL, main="EnsVoting",ylab="",xlab="", axis.col = NULL, axis.row = NULL,border="lightgrey")
msimilarity<-fsSimilarityMatrix(mlist, threshold = 0.1, method="Kuncheva")

needs(heatmaply)
plot(msimilarity, digits=2, col=rev(Spectral(21)),breaks=seq(-1,1,by=0.1), key=NULL, main="Sample-wise", cex.axis=0.7,ylab="",xlab="")
matrixes<-list()
for (f in list.files(CCF.folder.drift)){
  mlist<-list(
    CCF=readRDS(file.path(CCF.folder.drift,f)),
    GLASSO=readRDS(file.path(GLASSO.folder.drift,f)),
    MI=readRDS(file.path(MI.folder.drift,f)),
    LARS=readRDS(file.path(LARS.folder.drift,f)),
    RF=readRDS(file.path(RF.folder.drift,f)),
    Distance=readRDS(distance.folder),
    EnsVoting=readRDS(file.path(EnsVoting.folder.drift,f)),
    EnsRanking=readRDS(file.path(EnsRanking.folder.drift,f))
  )
  print(f)
  matrixes[[length(matrixes)+1]]<-fsSimilarityMatrix(mlist, threshold = 0.1, method="Kuncheva")
}
res<-apply(simplify2array(matrixes), 1:2, mean)
plot(res, digits=2, col=rev(Spectral(21)),breaks=seq(-1,1,by=0.1), key=NULL, main="Dynamic with drift", cex.axis=0.7,ylab="",xlab="")


fmt <- "%Y-%m-%d_%H%M%S"
day<-as.POSIXct("2017-08-02_000000",fmt)
mlist<-loadDay(day,CCFg.folder)
res<-fsSimilarityMatrix(mlist, threshold = 0.1,method="Kuncheva")
r<-res[order(rownames(res)), order(colnames(res))]
r
image(r,axes=FALSE)
```

Estimate complexity of feature sets
```{r complexity}
install.packages("graph")
require(needs)
BiocManager::install("Rgraphviz")
needs(QuACN)
needs(Rgraphviz)
set.seed(123)
g <- randomEGraph(as.character(1:10), 0.6)
offdiagonal(g)


convertToGraph<-function(fs){
  mEmpty <- matrix(0, nrow=nrow(fs),ncol=ncol(fs))
  adjM<-cbind(fs,mEmpty,mEmpty)
  adjM<-adjM+t(adjM)
  adjM[adjM>0]<-1
  colnames(adjM)<-rownames(adjM)
  g<-as(adjM,"graphNEL")
  return(g)
}

for (f in c(distance.folder,CCF.global, GLASSO.global, MI.global,LARS.global,RF.global,EnsVoting.global,EnsRanking.global)){
  print(paste(f,offdiagonal(convertToGraph(cutoff(readRDS(f),threshold=0.1)))))
}

options(pillar.sigfig = 4)

res<-list()
for (fold in c(CCFg.folder,MIg.folder,GLASSOg.folder,LARSg.folder,RFg.folder,EnsVotingg.folder,EnsRankingg.folder)){
  for (f in list.files(fold)){
    mFS=readRDS(file.path(fold,f))
    m<-cutoff(mFS,threshold=0.1)
    g<-convertToGraph(m)
    val <- offdiagonal(g)
    print(paste(fold,f,val))
    res[[length(res)+1]]<-list(folder=fold,file=f,odc=val)
  }
}
all<-bind_rows(res)
all%>%group_by(folder)%>%summarise(mean.odc = mean(odc))

```

Forecasting
```{r models}
RunModels <- function(data, series, ta, forecastingSteps, forecastEvery,
                      validationSize, validationEnd, 
                      lagMatrix,shortestA, cvgrid, results.folder){
  dir.create(results.folder, recursive = T)
  start.time <- Sys.time()
  results<-tibble()
  models.estimated <- c()
  if (file.exists(models.estimated.file)) models.estimated <- readRDS(models.estimated.file)
  else saveRDS(models.estimated, models.estimated.file)
  if (file.exists(results.file)) results <- readRDS(results.file)
  else saveRDS(results, results.file)
  for (r in 1:nrow(cvgrid)){
    cv <- cvgrid[r,]
    print(paste("Tuned parameter's set",r,"of",nrow(cvgrid)))
    print(cv)
    validationStart <- validationEnd - validationSize
    trainingSize <- cv$trainingMinutes/ta
    dat.restricted<-data[(validationStart-trainingSize+1):validationEnd,]
    base_params<-list(data=dat.restricted, seriesNames=series,
                      forecastingSteps=forecastingSteps, forecastEvery=forecastEvery)
    params<-c(base_params,list(trainingWindowSize=trainingSize))
    
    if (length(forecastingSteps)>1) params$forecastingSteps=max(forecastingSteps)
    #estimate.HA(params,cv,results.file, models.estimated.file)
    #estimate.naive(params,cv,results.file, models.estimated.file)
    #estimate.arima(params,cv,results.file, models.estimated.file)
    estimate.SpVAR(params,cv,results.folder)
    estimate.KNN(params,cv,results.folder)
    print(Sys.time() - start.time)
  }
}
```

Forecasting - experiments
```{r experiments}
dat<-as.data.frame(sampl$data)
rownames(dat)<-format(dat$datetime,"%Y-%m-%d %H:%M:%S")

# source(file.path("R","models_spvar.R"))
# xModel.star$run(dat[1:1440,series], forecastingSteps=3, arLags=3,
#                 fs.folder=RFg.folder,
#                 control=list(),
#                 returnModel=F,
#                 refine = F,
#                 include.mean=F,
#                 verbose=T,
#                 exclude.series=NULL,
#                 threshold=0.1)


week<-24*60*7
data<-sampl$data
validationSize <- nrow(data)-week/ta
validationEnd <- nrow(data)
forecastEvery<-4
forecastingSteps<-12

# Define a grid for tuning parameters
cvgrid <- expand.grid(trainingMinutes = c(week),
                      include.mean = c(F),
                      arLags = c(maxlag),
                      stationary =  c(T),
                      allowdrift =  c(F),
                      fs.folder=c(EnsVoting.folder.drift,
                                  EnsRanking.folder.drift),
                      threshold=c(0.1),
                      ownlags=c(T),
                      kneighbours=c(36))
source(file.path("R","models_spvar.R"))
source(file.path("R","cv_functions.R"))
# Run all models
RunModels(data, series, ta, forecastingSteps, forecastEvery,
                      validationSize, validationEnd, 
                      lagMatrix, shortestA, cvgrid,
                      results.folder="thtrain-hourly")
 
```


```{r analyse_results}
#params <- c('trainingMinutes','allowdrift','stationary','arLags','include.mean','fs.folder','threshold','ownlags')
# models.estimated <- readRDS(file.path(dir,"baseline.models.rds"))
# results <- readRDS(results.file)
# models.estimated <- readRDS(models.estimated.file)
# results<-results%>%mutate(fs.folder=gsub(file.path(dir,"fs",64),"",fs.folder))
# summary(results$actual)
# #results<-results%>%filter(abs(actual)>100)
# #results%>%select(datetime,detector,actual,forecasted)
# actu<-mysample$data.orig%>%gather(key="detector", value="actual.orig", -one_of("datetime"))%>%
#   mutate(detector=gsub(".volume","",detector))
# r<-results%>%left_join(actu, by=c("datetime"="datetime","detector"="detector"))
# r1<-r%>%mutate(actual=actual+actual.orig,forecasted=forecasted+actual.orig)

reslist<-list()
results<-NULL
fold<-file.path(dir,"thtrain-hourly","complete")
for (f in list.files(fold)){
  print(f)
  reslist[[length(reslist)+1]]<-readRDS(file.path(fold, f))
}
results<-bind_rows(reslist)
results<-readRDS(file.path(dir,"baseline.results.rds"))
params <- c('trainingMinutes','arLags','fs.folder','threshold','ownlags')
s<-cvSummary(results, rmse, cumulative=F, params=params)%>%
    mutate(modelid=paste(model,trainingMinutes, arLags,fs.folder,threshold,ownlags))%>%
    mutate(modelid=gsub(" NA","",modelid))
s%>%ungroup%>%mutate(key=paste(model,trainingMinutes,fs.folder))%>%select(key,threshold,forecast_horizon,RMSE)%>%spread(key=forecast_horizon, value=RMSE)%>%print(n=34)

```

Plots for forecasting accuracy
```{r forecasting_accuracy_charts}
gtmp<-results%>%filter(forecast_horizon==3)%>%mutate(key=paste(model,fs.folder))%>%group_by(key,datetime)%>%summarise(mn=rmse(actual,forecasted))
gtmp<-gtmp%>%ungroup%>%mutate(key=ifelse(str_detect(key,"CCF"),"CCF",key))%>%
    mutate(key=ifelse(str_detect(key,"GLASSO"),"GLASSO",key))%>%
    mutate(key=ifelse(str_detect(key,"MI"),"MI",key))%>%
    mutate(key=ifelse(str_detect(key,"LARS"),"LARS",key))%>%
    mutate(key=ifelse(str_detect(key,"RF"),"RF",key))%>%
    mutate(key=ifelse(str_detect(key,"EnsVoting"),"EnsVoting",key))%>%
    mutate(key=ifelse(str_detect(key,"EnsRanking"),"EnsRanking",key))
gtmp<-gtmp%>%filter(wday(datetime)==3)%>%mutate(dow_time=format(datetime, "%H"))%>%group_by(key,dow_time)%>%summarise(mn=mean(mn), last_date=last(datetime))

gtmp<-gtmp%>%spread(key=key,value=mn)%>%
    mutate(`CCF vs. EnsRanking`=CCF-EnsRanking)%>%
    mutate(`MI vs. EnsRanking`=MI-EnsRanking)%>%
    mutate(`LARS vs. EnsRanking`=LARS-EnsRanking)%>%
    mutate(`RF vs. EnsRanking`=RF-EnsRanking)%>%
    mutate(`GLASSO vs. EnsRanking`=GLASSO-EnsRanking)%>%
    mutate(`EnsVoting vs. EnsRanking`=EnsVoting-EnsRanking)%>%
    select(last_date,`CCF vs. EnsRanking`,`MI vs. EnsRanking`,`LARS vs. EnsRanking`,`RF vs. EnsRanking`,
           `GLASSO vs. EnsRanking`)%>%
    gather(`CCF vs. EnsRanking`,`MI vs. EnsRanking`,`LARS vs. EnsRanking`,`RF vs. EnsRanking`,
           `GLASSO vs. EnsRanking`,key="key", value="Difference")

gtmp%>%ggplot(aes(x = last_date, y=Difference, group=key, col=key))+
  geom_ma(size=1,n=2,linetype=1)+
  scale_x_datetime(labels = date_format("%H:%M", tz="Europe/Helsinki"),date_breaks = "4 hours",expand = c(0,0))+
  xlab("Time of the day") + ylab("Difference in RMSE")+ guides(col=guide_legend(title="Pair-wise comparison"))+
  theme(legend.position="bottom")+scale_color_brewer(palette="Dark2")

gtmp<-results%>%filter(forecast_horizon==3)%>%mutate(key=paste(model,fs.folder))%>%group_by(key,datetime)%>%summarise(mn=rmse(actual,forecasted))
gtmp<-gtmp%>%ungroup%>%mutate(key=ifelse(str_detect(key,"CCFdrift"),"Drift",key))%>%
  mutate(key=ifelse(str_detect(key,"CCFg"),"Daytime",key))%>%
  mutate(key=ifelse(str_detect(key,"CCF/36"),"Dynamic",key))%>%
  mutate(key=ifelse(str_detect(key,"CCF/201"),"Global",key))
    
gtmp<-gtmp%>%filter(wday(datetime)==3)%>%mutate(dow_time=format(datetime, "%H:%M"))%>%group_by(key,dow_time)%>%summarise(mn=mean(mn), last_date=last(datetime))

gtmp<-gtmp%>%spread(key=key,value=mn)%>%
    mutate(`Global vs. Drift`=Global-Drift)%>%
    mutate(`Dynamic vs. Drift`=Dynamic-Drift)%>%
    mutate(`Daytime vs. Drift`=Daytime-Drift)%>%
    select(last_date,`Global vs. Drift`,`Dynamic vs. Drift`,`Daytime vs. Drift`)%>%
    gather(`Global vs. Drift`,`Dynamic vs. Drift`,`Daytime vs. Drift`,key="key", value="Difference")
  
gtmp%>%ggplot(aes(x = last_date, y=Difference, group=key, col=key))+
  geom_line(size=1,linetype=1)+
  scale_x_datetime(labels = date_format("%H:%M", tz="Europe/Helsinki"),date_breaks = "4 hours",expand = c(0,0))+
  xlab("Time of the day") + ylab("Difference in RMSE")+ guides(col=guide_legend(title="Pair-wise comparison"))+
  theme(legend.position="bottom")+scale_color_brewer(palette="Dark2")
```
